---
title: "PS270 Final Project"
author: "Holly Siemers"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to...misinformation on social media and how it effects political polarization. Does the misinformation increase it? Or how protests affect policy change.

## Project Milestone 2
```{r}
library(tidyverse)
library(readr)
library(stringr)
```

```{r}
wnp <- read_csv("World Bank Women in National Parliaments - Sheet1.csv")
wnp
```
```{r}
democracy_index <- read_csv("democracy-index-eiu/democracy-index-eiu.csv")
democracy_index
```

Is there a positive relationship between a country’s level of democracy and the percentage of women in its national legislature? In his study I will examine the correlation between countries with democracy and how many women are in office nationally in those countries. My hypothesis is: Countries with higher levels of democracy will have higher percentages of women in parliament. My explanatory variable is democracy level. I will be using the Our World in Data study for this. They measure democracy on a 0-10 scale, with higher values being more democratic. My outcome variable is the percentage of women in national legislatures, which I will find with the World Bank data on women in national legislatures. They measure the percentage of women by dividing the number of women seats by the total seats and multiplying it by 100. (it is a 0-100 scale). A pattern that would show support for my hypothesis would be a positive correlation between democracy and women in office. Some things that would show disapproval of my hypothesis would be if there was no statistical correlation between the two variables or if democratic countries show low % of women in office.

## Milestone 3

Was deleted. 

## Milestone 4

In this milestone I need to compile all my data together now that I have both needed data sets. I will focus on the most recent year in each data set which is 2024. 

```{r}
wnp <- wnp |> rename(Country = Economy)
democracy_index <- democracy_index |> rename(Country = Entity)
```


```{r}
democracy_wnp <- left_join(wnp, democracy_index, by = c("Country", "Year"))
democracy_wnp
```


```{r}
democracy_wnp <- na.omit(democracy_wnp)
democracy_wnp
```
```{r}
democracy_wnp <- democracy_wnp |> 
  rename(Prop_Seats = "Proportion of seats held by women in national parliaments (%)", Democracy_Index = "Democracy Index")
democracy_wnp
```
```{r}
library(dplyr)
democracy_wnp <- democracy_wnp |>
  filter(Year == 2024)
```



## Milestone 4 Continued
```{r}
 democracy_wnp <- 
  democracy_wnp |>
  filter(
    !is.na(Democracy_Index),
    !is.na(Prop_Seats)
  ) |>
  mutate(Prop_Seats = as.numeric(str_remove(Prop_Seats, "%")))
democracy_wnp
```

```{r, eval = F}
library(ggplot2)

ggplot(democracy_wnp, aes(x = Democracy_Index, y = Prop_Seats)) +
  geom_point(alpha = 0.6) +

  labs(
    title = "Democracy and Women's Representation in National Legislatures",
    subtitle = "Each point represents a country-year observation",
    x = "Democracy Score",
    y = "Women in Parliament (%)",
    caption = "Source: Our World in Data & World Bank"
  )
```
I need to fix this graph a lot. I need to change my data set section, Prop_Seats, into a numeric area. Right now I think it is reading differently and it is messing my graph up. Unfortunately I have spent a lot of time trying to put a large data set into R, trying to find a different one after it did not work properly, then compiling the new data sets together. I hope that after going to office hours this next week, I can fix these changes with extra help and be back on track. I apologize for not properly completing this weeks milestone.
